---
title: "Using `latexTable()` with R Markdown documents"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style>
  <!--

  a, a[href^="https:"] {
    text-decoration: none;
  }
  body {
    font-family: Georgia, "Times New Roman", Times, serif;
    font-size: 1.05em;
    line-height: 1.7;
    margin-top: 9ex;
    margin-bottom: 12ex;
  }
  div.sourceCode {
    margin-top: -1ex;
    margin-bottom: 1.5ex;
    margin-left: 0;
    margin-right: 0
  }
  h1 {
    margin-top: 3ex;
    margin-bottom: 0.67ex;
  }
  h2 { 
    border-bottom: unset; 
    /* margin-block-end: unset; */
    margin-bottom: 0.1ex;
    padding-top: 2.75ex;
  }
  p {
    margin: unset;
    margin-bottom: 2ex;
    max-width: 60ch;
  }
  
  /* For images immediately preceded by code chunks */
  .sourceCode + p > img {  
    margin-top: -4ex;
    margin-bottom: 4ex;
  }
  
  .marginBottom1p5ex { margin-bottom:  1.5ex; }
  .marginBottom2ex   { margin-bottom:  2.0ex; }
  .marginBottom4ex   { margin-bottom:  4.0ex; }
  .marginTopN4p5ex   { margin-top:    -4.5ex; }

  /* Fix display of source code on small screens. See #25. */
  a.sourceLine { display: unset; }
  pre > code.sourceCode > span { display: unset;  }

  -->
</style>


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


It is easy to use `latexTable()` in R Markdown documents, provided that 
you are generating LaTeX or PDF files from those R Markdown documents. To do
so, you need to take only two steps: load a file via the `in_header` option 
for R Markdown documents, and wrap your `latexTable()` output in the 
`raw_latex()` command that is provided by the `knitr` package.



## 1. Header
To use `latexTable()` in R Markdown documents, certain LaTeX packages must be 
loaded: the `afterpage, array, booktabs, caption, numprint` and `ragged2e` 
packages. Depending on the options that you specify in your `latexTable()` 
call, you may also need to load the `afterpage`, `float`, and `pdflscape`
packages. 

The `Bullock` package comes with a file, `latexTable_RMarkdown_preamble.tex`,
that loads all of these packages. Thus, to load the packages, you need only 
tell R Markdown to load `latexTable_RMarkdown_preamble.tex`. The best way to 
do this is by using the `in_header` option in the header of your R Markdown 
document. The header should look like this:

    ---
    title: "Test latexTable() Output"
    output: 
      pdf_document:
        includes:
          in_header: !expr system.file("latexTable_RMarkdown_preamble.tex", package = "Bullock", mustWork = TRUE)
    ---



## 2. Inserting `latexTable()` output 
To insert `latexTable()` output into your document,  just wrap the output in 
a `raw_latex()` command:

```{r}
library(Bullock)
library(knitr)
myTable <- latexTable(matrix(1:16, nrow = 4))
raw_latex(myTable)
```
Remember that if you are using multiple tables, you will need to specify 
a unique LaTeX name for each one (by using the `commandName` argument in your 
`latexTable()` calls).



## 3. Limitations
The most obvious limitation is that `latexTable()` cannot be used when 
generating HTML documents from R Markdown. It will work only when generating 
LaTeX or PDF from R Markdown.